---
title: "challenge02"
author: "Jyhreh Johnson"
date: "2/24/2022"
output: html_document
---
Challenge 2

1. Using the {tidyverse} read_csv() function, load the “zombies.csv” dataset from this URL as a “tibble” named z.
```{r}
library(tidyverse)
f1 <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/zombies.csv"
z <- read_csv(f1, col_names = TRUE)
head(z)
```
.
.
.
2. Calculate the population mean and standard deviation for each quantitative random variable in the dataset (height, weight, age, number of zombies killed, and years of education).
```{r}
library(radiant)
p_means <- dplyr::summarize(z, mean(height), mean(weight),  
                            mean(age), mean(zombies_killed),
                            mean(years_of_education)) 

p_sd <- dplyr::summarize(z, sdpop(height), sdpop(weight),  
                            sdpop(age), sdpop(zombies_killed),
                            sdpop(years_of_education))
```
.
.
.
3. Use {ggplot} and make boxplots of each of these variables by gender.
```{r}
library(cowplot)
#gender x height
zb1 <- ggplot(data = z, aes(x = gender, y = height))
zb1 <- zb1 + geom_boxplot(na.rm = TRUE)
zb1 <- zb1 + theme(axis.text.x=element_text(angle = 90))
zb1 <- zb1 + xlab("gender") + ylab("height")

#gender x weight
zb2 <- ggplot(data = z, aes(x = gender, y = weight))
zb2 <- zb2 + geom_boxplot(na.rm = TRUE)
zb2 <- zb2 + theme(axis.text.x=element_text(angle = 90))
zb2 <- zb2 + xlab("gender") + ylab("weight")

#gender x age
zb3 <- ggplot(data = z, aes(x = gender, y = age))
zb3 <- zb3 + geom_boxplot(na.rm = TRUE)
zb3 <- zb3 + theme(axis.text.x=element_text(angle= 90))
zb3 <- zb3 + xlab("gender") + ylab("age")

#gender x zombies_killed
zb4 <- ggplot(data = z, aes(x = gender, y = zombies_killed))
zb4 <- zb4 + geom_boxplot(na.rm = TRUE)
zb4 <- zb4 + theme(axis.text.x=element_text(angle = 90))
zb4 <- zb4 + xlab("gender") + ylab("zombies_killed")

#gender x years_of_education
zb5 <- ggplot(data = z, aes(x = gender, y = years_of_education))
zb5 <- zb5 + geom_boxplot(na.rm = TRUE)
zb5 <- zb5 + theme(axis.text.x=element_text(angle = 90))
zb5 <- zb5 + xlab("gender") + ylab("years_of_education")

plot_grid(zb1, zb2, zb3, zb4, zb5, label_size = 12, nrow =2)
```
.
.
.
4. Use {ggplot} and make scatterplots of height and weight in relation to age (i.e., use age as the x variable), using different colored points for males versus females.
```{r}
#age x height
z_scat1 <- ggplot(data = z, aes(x = age, y = height, color = factor(gender)))
z_scat1 <- z_scat1 + xlab("age") + ylab("height")
z_scat1 <- z_scat1 + geom_point(na.rm = TRUE)
z_scat1 <- z_scat1 + theme(legend.position = "bottom", legend.title = element_blank())

#age x weight
z_scat2 <- ggplot(data = z, aes(x = age, y = weight, color = factor(gender)))
z_scat2 <- z_scat2 + xlab("age") + ylab("weight")
z_scat2 <- z_scat2 + geom_point(na.rm = TRUE)
z_scat2 <- z_scat2 + theme(legend.position = "bottom", legend.title = element_blank())

plot_grid(z_scat1, z_scat2,label_size = 12, nrow = 2)
```
Do these variables seem to be related? In what way?
-Yes, height & age and weight & age appear to have linear relationships.
.
.
.
5. Using histograms and Q-Q plots, check whether each of the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not?
```{r}
#histograms
zq1 <- ggplot(data = z, aes(x  = height)) + geom_histogram(bins = 30) + ggtitle("Height")

zq2 <- ggplot(data = z, aes(x  = weight)) + geom_histogram(bins = 30) + ggtitle("Weight")

zq3 <- ggplot(data = z, aes(x  = age)) + geom_histogram(bins = 30) + ggtitle("Age")

zq4 <- ggplot(data = z, aes(x  = zombies_killed)) + geom_histogram(bins = 30) + ggtitle("Zombies_Killed")

zq5 <- ggplot(data = z, aes(x  = years_of_education)) + geom_histogram(bins = 30) + ggtitle("Years_of_Education")

#qqplots, couldn't figure out how to work qqnrom & qqline so i decided to use ggpubr
#install.packages(ggpubr)
library(ggpubr)
zq6 <- ggqqplot(z$height, title = "Height")
zq7 <- ggqqplot(z$weight, title = "Weight")
zq8 <- ggqqplot(z$age, title = "Age")
zq9 <- ggqqplot(z$zombies_killed, title = "Zombies_killed")
zq10 <- ggqqplot(z$years_of_education, title = "Years_of_Education")

plot_grid(zq1, zq2, zq3, zq6, zq7, zq8, label_size = 12, nrow = 2)
plot_grid(zq4, zq5, zq9, zq10, label_size = 12, nrow = 2 )
```
Which seem to be and which do not?
-Height, Weight, and Age all appear to be evenly distributed, but Zombies_Killed
and Years_of_education are not normally distributed. ZK and YOE may be plotted 
with a different distribution???
.
.
.
6. Now use the sample_n() function from {dplyr} to sample ONE subset of 50 zombie apocalypse survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable based on this one sample and use that to construct the theoretical 95% confidence interval for each mean. You can use either the standard normal or a Student’s t distribution to derive the critical values needed to calculate the lower and upper limits of the CI.
```{r}
```
.
.
.
7. Then draw another 199 random samples of 50 zombie apocalypse survivors out and calculate the mean for each of the these samples. Together with the first sample you drew out, you now have a set of 100 means for each variable (each based on 50 observations), which constitutes a sampling distribution for each variable.
```{r}
```
What are the means and standard deviations of the sampling distribution for each variable?
-
.
.
.
8. Finally, construct a 95% confidence interval for each mean directly from the sampling distribution of sample means using the central 95% that distribution (i.e., by setting the lower and upper CI bounds to 2.5% and 97.5% of the way through that distribution).
```{r}
```
.
.
.
9. How do the standard deviations of the sampling distribution for each variable compare to the standard errors estimated from your first sample of size 50? 
-
.
.
What do sampling distributions for each variable mean look like? Are they normally distributed? 
-
.
.
What about for those variables that you concluded were not originally drawn from a normal distribution? 
-
.
.
How do the two 95% CIs you estimated compare to one another (i.e., the CI based on one sample and the corresponding sample standard deviation versus the CI based on simulation where you created a sampling distribution across 200 samples)?
-